---
title: "fastLink tutorial"
author: "Emanuel Ben-David"
date: "`r Sys.Date()`"
output:
  beamer_presentation: default
  slidy_presentation: default
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## What is fastLink?

-   **fastLink** is an **R** package for Fast Probabilistic Record Linkage

-   fastLink to some degree can handle missing data

-   It is amenable to parallel computing, using the number of cores in the
    computer

-   **splink** is essentially a translation of fastLink to python language, with
spark capability, for **python users**.

-   It is written by **Ted Enamorado** [aut, cre], Ben Fifield [aut], Kosuke Imai

## Linking two files

We would like to link two data files: dfA and dfB. The main function for linking
is fastlink().

fastLink(dfA, dfB, varnames, stringdist.match, stringdist.method, numeric.match,
partial.match, cut.a, cut.p, ...)

-   "**varnames**": vector of matching variables.
-   Must be present in both dfA and dfB
-   "**stringdist.match**": vector of string variables in "varnames"
-   "**stringdist.method**": default is jw" for Jaro-Winkler, other options are 
"jaro" for Jaro, and "lv" for edit.

##   Linking files with fastlink() continued 

-   **"numeric.match": numeric** variables for numeric matching

-   **"partial.match":** string variables among "**stringdist.match" variables**
    for partial matching.
-   "**cut.a**": lower bound for full string-distance match, ranging between 0
     and 1. Default is 0.94
-   "**cut.p**": Lower bound for partial string-distance match, ranging between
    0 and 1. Default is 0.88
-   "**n.cores**": number of cores to parallelize over. Default is NULL.    
    
-  We can use "getMatches()" function to get the matches.

-  The arguments for "getMatches()" are:

-  dfA: files A
-  dfB : file B
-  fl.out: the output of "fastlink()" in the setp above.


    
## Slide with R Output

```{r}
library(fastLink)
help("fastLink")
```

-  For demo, we uplaod file_a and file_b.

```{r echo = F, include=FALSE}
library(readxl)
file_a <- read_excel("/home/emanuelben-david/IISA_short_course/data/file_a.xlsx")

file_b <- read_excel("/home/emanuelben-david/IISA_short_course/data/file_a.xlsx")

```

` Let's see the variables in data files.

```{r}
colnames(file_a)
colnames(file_b)
```


-   We link these two data files by blocking on
"BirthYear".



```{r echo=TRUE}
library(dplyr)
file_a_1850 <- file_a%>%filter(BirthYear == 1850)
file_b_1850 <- file_b%>%filter(BirthYear == 1850)

```

- For this demo, we just link two blocks for "BirthYear" == 1850

` Let's check the size of these blocks ("BirthYear" == 1850)

```{r}
file_a_1850%>%dim()
file_b_1850%>%dim()

```

- For linking with fastLink, we set:
```{r}
varnames = c("LastName", "FirstName", "MiddleName")
stringdist.match = c("LastName", "FirstName", "MiddleName")
partial.match = "MiddleName"

```

-  We choose the DEFAULT values for the rest of fastLink arguments.

-  Now we apply fastlink()

```{r}
 
fastlink_out <- fastLink(file_a_1850, file_b_1850,
                         varnames = varnames, stringdist.match = stringdist.match, partial.match = partial.match)

```

` Using getMatches() we obtain the mathces in each files.

```{r echo=TRUE}

matches_ab_1850 <- getMatches(file_a_1850, file_b_1850, fastlink_out,
                            combine.dfs = FALSE)
  
```


- We check the matches:

```{r}

matches_ab_1850%>%head(4)
matches_ab_1850%>%dim()

```
